---
title: "results experiment 1"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggforce)
library(here)
library(tidyverse)
source(here("R", "utils.R"))

# data_dir <-  here("data", "test-runs")
data_dir <- here("data", "prolific")

# result_dir <- here("data", "test-runs", "experiment1")
result_dir <- here("data", "prolific", "results", "experiment1-v2")

fn <- "exp1_tidy.rds"
# fn <- "exp1_tidy_quality.rds"

data.from_rds <- readRDS(paste(result_dir, fn, sep=.Platform$file.sep))
if(str_split(fn, "_")[[1]] %>% length() == 2) {
  data <- data.from_rds$test
} else {
  data <- data.from_rds
}

data.orig <- data %>% mutate(response=r_orig) %>% select(-r_norm, -r_orig)
data.norm <- data %>% mutate(response=r_norm) %>% select(-r_norm, -r_orig)
# choose wether normalized or unnormalized values shall be plotted
data.plot <- data.norm


stimuli <- data.orig %>% pull(id) %>% unique()
N_participants <- data.orig %>% pull(prolific_id) %>% unique() %>% length()

plot_dir <- paste(result_dir, "plots", sep=.Platform$file.sep);
dir.create(plot_dir, showWarnings = FALSE)
```

Use normalized data and check how many all-or-none answers were given since we
dont really want them, stimuli should have a little more uncertainty than that.

```{r, eval=FALSE}
# Look for participants who gave all-or-none responses, if any,
# which stimuli were rated as all-or-none by how many participants?
df <- data.norm %>% group_by(prolific_id, id) %>%
  summarise(n=max(response)) %>%
  filter(n==1)

N0 <- N_participants - df %>% pull(prolific_id) %>% unique() %>% length()
df0 <- tibble(prolific_id=rep('--', N0), ratio_all_none=rep(0, N0))

df.participants_all_none <- df %>% group_by(prolific_id) %>%
  summarise(ratio_all_none = n() / length(stimuli)) %>%
  mutate(prolific_id=as.character(prolific_id)) %>%
  bind_rows(df0)

if (nrow(df.participants_all_none) != 0){
  p <- df.participants_all_none %>%
        ggplot(aes(ratio_all_none)) + geom_bar() +
        labs(y="#participants", x="#all-none answers")
  print(p)
  print(df.participants_all_none)

  df.stimuli_all_none <- df %>%
    ungroup() %>% group_by(id) %>% summarise(n_participants = n())

  p <- df.stimuli_all_none %>%
        ggplot(aes(x=id, y=n_participants)) + geom_bar(stat="identity") +
        theme(axis.text.x = element_text(angle=90, hjust=1)) +
        labs(y="#participants who gave all-none answer")
  print(p)
} else {
  print('nobody gave all-none responses (where sum=1)')
}
```


## Experimental data: probability tables

```{r}
dat <- data.plot %>% ungroup() %>%
  mutate(question = factor(question, levels = c("bg", "b", "g", "none")))

labels <- c(bg="Blue, Green", b="Blue, ¬Green",
            g="¬Blue, Green", none="¬Blue, ¬Green")

df <- cluster_responses(dat, "g") %>%
  group_by(question, id) %>% mutate(mu=mean(response), med=median(response));

n_pages <- 14/2 # 2 test trials per page!
for (pa in seq(1, n_pages)){
  p <- df  %>%
    ggplot(aes(x=response, y=factor(0))) +
    geom_jitter(height = 0.1, alpha=0.5, aes(shape = cluster, colour = prolific_id)) +
    geom_point(mapping=aes(x=mu, y=factor(0)),  shape=8, color = "red") +
    # geom_vline(aes(xintercept=med), color = "red", size=0.5) +
    geom_vline(aes(xintercept=0), color = "grey", size=0.2, linetype='dashed') +
    geom_vline(aes(xintercept=0.25), color = "grey", size=0.2, linetype='dashed') +
    geom_vline(aes(xintercept=0.5), color = "gray45", size=0.5, linetype='dashed') +
    geom_vline(aes(xintercept=0.75), color = "grey", size=0.2, linetype='dashed') +
    geom_vline(aes(xintercept=1), color = "grey", size=0.2, linetype='dashed') +
    scale_x_continuous(limits=c(-0.2, 1.2),
                       breaks=c(0, 0.25, 0.5, 0.75, 1)) +
    labs(x="estimates probability", y="") +
    theme_bw() +
    facet_grid_paginate(id~question, nrow=2, ncol = 4,
                        labeller = labeller(question=labels),
                        page = pa) +
    theme(axis.text.y=element_blank(),
          axis.ticks.y =element_blank(),
          text = element_text(size=14),
          axis.text.x=element_text(angle=45, vjust = 0.5),
          legend.position = "right")
  print(p)
}
```


## Sorted according to expectations
Sort trials according to how they are expected to be interpreted
(depending on which utterances we expect)


```{r}
df <- data.norm %>%
  pivot_wider(names_from = question, values_from = response) %>%
  ungroup() %>% mutate(id=as.character(id)) %>% add_probs() %>%
  select(-group, -QUD, -n)
as_independent <- c('if1_hh', 'if1_lh',
                    'if2_ll', 'if2_hh', 'if2_uh')
ind <- df %>% filter(startsWith(id, "independent") | id %in% as_independent) %>%
  add_column(kind = "independent")

as_a_implies_c_cause <- c('if2_hl', 'if2_ul',
                          'if1_uu', 'if1_uh', 'if1_hh')
ac_cause <- df %>% filter(id %in% as_a_implies_c_cause) %>%
  add_column(kind = "cause")

as_a_implies_c_diagnostic <- c('if2_ul', 'if1_uu')
ac_diagnostic <- df %>% filter(id %in% as_a_implies_c_diagnostic) %>%
  add_column(kind = "diagnostic")

others <- df %>% filter(id %in% c('if1_uu')) %>%
  add_column(kind = "other")

# N <- ind %>% nrow() + others %>% nrow() + ac_cause %>% nrow() +
#   ac_diagnostic %>% nrow()
# stopifnot(N == nrow(df))
```

prepare the data

```{r, echo=FALSE}
dataIndependent = function(data) {
  stimuli <- data %>% pull(id) %>% unique()
  df <- data %>% mutate(prod_ac=p_a*p_c, p=bg-prod_ac, p_key = "diff_ac_prod_ac") %>%
    select(prolific_id, id, p, p_key)

  labels <- c(diff_ac_prod_ac="P(B,G)-P(B)xP(G)")
              # , bg="P(Blue,Green)", prod_ac="P(Blue) * P(Green)")
  return(list(labels=labels, data=df, stimuli=stimuli))
}

dataImplies <- function(data){
  stimuli <- data %>% pull(id) %>% unique()
  df <- data %>%
    select(prolific_id, id, p_c, p_c_given_a) %>%
    mutate(p = p_c_given_a - p_c, p_key = factor("delta_p")) %>%  
    select(-p_c, -p_c_given_a)
  labels <- c(delta_p="P(C|A)-P(C)")
  return(list(labels=labels, data=df, stimuli=stimuli))
}

dataOthers <- function(data){
  stimuli <- data %>% pull(id) %>% unique()
  df <- data %>%
    select(prolific_id, id, p_c, p_c_given_a, none) %>%
    mutate(p_delta = p_c_given_a - p_c, p_delta_star = if_else(none==0, 1, p_delta/none)) %>%
    select(prolific_id, id, p_delta_star) %>%
    rename(p=p_delta_star) %>%
    add_column(p_key=factor("delta_p_star"))
  labels <- c(delta_p_star="(P(C|A)-P(C))/P(¬A,¬C)")

  return(list(labels=labels, data=df, stimuli=stimuli))
}
```

Plot function

```{r}
plotProbs <- function(df, xlab, title, limits_x = c(-0.2, 1.2),
                      breaks_x = c(0, 0.25, 0.5, 0.75, 1)) {
  df_means <- df$data %>% as_tibble() %>% group_by(id, p_key) %>%
                summarise(mu=mean(p), med=median(p))
  dat <- left_join(df$data, df_means)
  n = 1; pa = 1;
  while(pa <= n) {
    p <- dat %>%
      ggplot(aes(x=p, y=factor(0))) +
      geom_jitter(height = 0.1, alpha=0.5, aes(colour=factor(prolific_id))) +
      scale_x_continuous(limits=limits_x, breaks=breaks_x) +
      labs(x=xlab, y="") +
      theme_bw() +
      facet_wrap_paginate(c("id"), nrow=2, ncol = 4, page = pa) +
      geom_vline(aes(xintercept=med), color="orange", size=0.5) +
      geom_point(aes(x=mu, y=factor(0)), color="red", shape=24, size=0.75) +
      theme(legend.position = "right",
            axis.text.y=element_blank(),
            axis.ticks.y =element_blank(),
            axis.text = element_text(size=11, angle=45, hjust=1),
            panel.spacing = unit(2, "lines")) +
      ggtitle(title)
    print(p)
    n <- n_pages(p)
    pa <- pa + 1
    # fn <- paste(plot_dir, paste("exp1-", s, "-probs.png", sep=""), sep=.Platform$file.sep)
    # ggsave(fn, p, width=8, height=5)
  }
}
```


## Trials that are expected to be INDEPENDENT
Here, P(A,C) - P(A) x P(C) should be close to zero.

```{r, warning=FALSE}
df <- dataIndependent(ind)
plotProbs(df, "P(A,C)-P(A)xP(C)", "expected interpretation: independent",
          limits_x = c(-0.5, 0.5), breaks_x = seq(-0.5, 0.5, 0.25))
```

## Trials that are expected to be A->C (cause).
Check delta p = P(C|A) - P(C).

```{r}
df_ifs <- bind_rows(ac_cause, ac_diagnostic)
df <- dataImplies(df_ifs)
min <- df$data$p %>% min() - 0.1
max <-  df$data$p %>% max() + 0.1
plotProbs(df, "delta p", " P(C|A)-P(C)",
          limits_x = c(min, max),
          breaks_x = c(seq(round(min, 1), round(max, 1) + 0.05, 0.3)))
```

Other trials

```{r}
df <- dataOthers(others)
min <- df$data$p %>% min() - 0.1
max <-  df$data$p %>% max() + 0.1
plotProbs(df, "delta P*", "delta P* = (P(C|A)-P(C))/(P(¬A,¬C)",
          limits_x = c(min, max),
          breaks_x = c(seq(round(min, 1), round(max, 1) + 0.05, 1)))
```


```{r, echo = FALSE, message=FALSE}
fn <- paste(result_dir, "exp1_tables_mean.csv", sep=.Platform$file.sep)
tables <- read_csv(fn) %>% add_probs() %>%
  pivot_longer(cols=starts_with("p_"), names_to = "key", values_to = "prob")
```


```{r, include=FALSE, eval=FALSE}
PlotDensity <- function(dat, p1, p2, save_as){
  p <- dat %>%
    filter(key==p1 | key == p2) %>%
    ggplot(aes(prob, fill=key)) +
    geom_density(alpha=0.5)

  # ggsave(paste(TARGET_DIR, save_as, sep=.Platform$file.sep), p, width=6, height=4)
  return(p)
}
```


```{r, echo=FALSE, eval=FALSE, include=FALSE}
# Distributions of peoples' beliefs about probabilities across all stimuli

tables %>% PlotDensity("p_a_given_c", "p_c_given_a", "p-conditionals.png")
tables %>% PlotDensity("p_a", "p_c", "p-marginals.png")
tables %>% PlotDensity("p_na_given_nc", "p_nc_given_na", "p-neg-conditionals.png")
tables %>% PlotDensity("p_na", "p_nc", "p-neg-marginals.png")

```
