---
title: "stats Experiment 1"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggforce)
library(here)
library(scales)
library(tidyverse)
source(here("R", "utils.R"));

# result_dir <-  here("data", "test-runs");
result_dir <- here("data", "prolific", "results", "experiment1")
data_fn <- "exp1_tidy.rds"
path_data <- paste(result_dir, data_fn, sep=.Platform$file.sep)

data <- readRDS(path_data)
```

## Duration

```{r}
df <- data$info %>% group_by(prolific_id) %>% distinct()

p <- df %>%
  ggplot(aes(x=factor(0), y=timeSpent)) + geom_boxplot() + 
  geom_jitter(shape=16) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + sd(df$timeSpent),
                 color="yellow")) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + 2 * sd(df$timeSpent),
                 color="red")) +
  theme(legend.position="none") + 
  labs(x = "", y="timeSpent in min.")
p

df %>% ungroup() %>%  group_by(prolific_id) %>%  summary()
```


```{r, include=FALSE, eval=FALSE, echo=FALSE}
## Train Trials Color Vision

df <- data$color %>% group_by(prolific_id, id) %>%
  mutate(correct = expected == response)

df %>%
  ggplot(aes(x=id)) + geom_bar(position="dodge", aes(fill=correct)) +
  theme(axis.text.x = element_text(angle=90, hjust=1), legend.position = "top")

```

## Train Trials

```{r}
df <- data$train %>% group_by(prolific_id, id)
df.slider <- df %>% filter(trial_name == "animation_slider") 
df.buttons_correct = df %>%
  filter(case_when(expected == "response1" ~ icon_idx == 1,
                   expected == "response2" ~ icon_idx == 2, 
                   expected == "response3" ~ icon_idx == 3,
                   expected == "response4" ~ icon_idx == 4, 
                   ))
# not everyone saw the same trial as slider - trial, therefore sums will not 
# always be exactly the number of participants.
df.buttons_correct %>%
  ggplot(aes(x=id)) + geom_bar(position="dodge", aes(fill=response)) + 
  theme(axis.text.x=element_text(angle=90, hjust=1), legend.position = "top") + 
  guides(fill = guide_legend(title = "correct"))


```


## Reaction times (per experimental trial)

```{r}
df <- data$test %>% select(prolific_id, id, RT) %>% group_by(id) %>% distinct()

dat.RT = df %>% group_by(id) %>% summarise(mu=mean(RT), sd=sd(RT))
df <- left_join(df, dat.RT, by="id")
for(pa in seq(1,2)) {
  p <- df %>%  
    ggplot(aes(y=RT)) + geom_boxplot() +
    geom_jitter(aes(x=0, color=prolific_id), width = 0.1, alpha = 0.5) +
    geom_hline(aes(yintercept = mu), color="green", linetype="dashed") +
    theme(legend.position="none", axis.ticks.x = element_blank(), axis.text.x = element_blank()) +
    labs(x="", y="RT in ms", title="Reaction Times per stimulus") +
    facet_wrap_paginate(~id, nrow=2, ncol=5, page=pa) +
    scale_y_continuous(trans = log2_trans(),
                       breaks = trans_breaks("log2", function(x) 2^x),
                       labels = trans_format("log2", math_format(2^.x)))
  print(p)
}
```

# Quality of the data
For each participant and trial, compute the distance in RESPONSE to all other participants
and compute the average distance for this participant from all other participants. 

```{r, echo = FALSE, message = FALSE}
df <- data$test %>% unite(col = "id_quest", "id", "question", remove=FALSE) %>% 
  select(-r_norm) %>% rename(response=r_orig)
distances <- tibble()
for(proband in df$prolific_id %>% unique()) {
  message(proband)
  res = df %>% filter(prolific_id == proband) %>% ungroup()
  for(stimulus in df$id %>% unique()) {
    dat <- df %>% filter(id == stimulus);
    res_proband = res %>% filter(str_detect(string = id_quest, pattern = paste(stimulus, ".*", sep=""))) %>% 
      select(id_quest, response) %>% rename(r_proband = response) %>% 
      add_column(comparator = proband)
    distances <- bind_rows(distances,
                  dat %>% left_join(res_proband, by="id_quest") %>% mutate(diff = r_proband - response)); 
  }
}
distances <- distances %>% filter(comparator != prolific_id) %>%
  group_by(comparator, id, question) %>% mutate(m=mean(diff))
```

```{r}
facet_labels <- c(
  `bg` = "Blue and Green", 
  `b` = "Blue but not Green",
  `g` = "Green but not Blue",
  `none` = "Neither Green nor Blue"
);

df <- distances %>%
  select(comparator, question, m, id) %>%
 ungroup() %>% mutate(question = factor(question, levels=c("bg", "b", "g", "none"))) %>% 
 group_by(comparator, question) %>% unique() %>% ungroup()

stats_all <- df %>% group_by(question, id) %>% mutate(mu=mean(m), sd=sd(m)) %>% 
  select(question, id, mu, sd) %>% unique() %>% ungroup() %>%
  mutate(limit_up = mu + sd, limit_low = mu - sd)

df <- left_join(df, stats_all) %>% group_by(question, id)
for(pa in seq(1, 10)) {
  p <- df %>% 
     ggplot(aes(y=m, x=factor(0))) +
     geom_boxplot() +
     geom_jitter(aes(color=comparator), width = 0.1, alpha = 0.5) +
     geom_hline(aes(yintercept=mu), linetype="dashed", color = "orange", size=.5) +
     # geom_hline(aes(yintercept=limit_up), color = "black", size=.25) +
     # geom_hline(aes(yintercept=limit_low), color = "black", size=.25) +
     scale_y_continuous(limits=c(-1.1, 1.1), breaks=seq(-1, 1, 0.5)) +
     facet_grid_paginate(id~question, nrow = 2, ncol = 4,
                         labeller = labeller(question = facet_labels),
                         page = pa) + 
     ggtitle("Quality of responses per stimulus") +
     theme(axis.ticks.x = element_blank(), axis.text.x = element_blank(),
           legend.position="none") +
     labs(y="average difference to other participants' responses", x="")
  print(p)
}
```



General comments

```{r}
data$comments
data$comments$comments
data$comments$noticed_steepness
```

# Postprocessing

```{r}
filter_prolific_id <- function(df){
  df <- df %>%
    filter(prolific_id != "participant2" & prolific_id != "participant3" &
           prolific_id != "participant11" & prolific_id != "participant13" &
           prolific_id != "participant14")
  message(paste('# filtered out because of not noticing steepness:', "5"))
  return(df)
}

filter_quality <- function(df) {
  df <- left_join(df, stats_all) %>%
    filter(r_orig <= limit_up & r_orig >= limit_low) %>% 
    select(-limit_up, -limit_low, -mu, -sd)
  return(df)
}
```

##Filter out participants according to certain criteria
1. data of participants who did not notice different degrees of inclines
2. stimuli where average diff. to other participants' responses is smaller or
larger than one standard deviation

```{r}
update_data <- function(result_dir, func_filter, fn_filtered){
  target_fn <-"exp1_tidy.rds";
  target_path <- paste(result_dir, target_fn, sep=.Platform$file.sep);
  
  updated_fn <- paste("exp1_tidy_", fn_filtered, ".rds", sep="")
  df <- readRDS(target_path)$test %>% func_filter();
  saveRDS(df, paste(result_dir, updated_fn, sep=.Platform$file.sep))
  return(df)
}
  # also update (mean) tables
  # tables.all <- df %>% select(id, question, prolific_id, r_norm) %>%
  #   group_by(id, question, prolific_id) %>%
  #   pivot_wider(names_from = question, values_from = r_norm)
  # write.table(tables.all, file=paste(result_dir,
  #   "exp1_prob_tables_all_steepness_noticed.csv", sep=.Platform$file.sep),
  #   sep = ",", row.names=FALSE)
  # 
  # means <- df %>% group_by(id, question) %>% summarise(mean=mean(r_norm))
  # write.table(means %>% pivot_wider(names_from = question, values_from = mean),
  #             file=paste(result_dir, "exp1_prob_tables_mean_steepness_noticed.csv", sep=.Platform$file.sep),
  #             sep = ",", row.names=FALSE)

df1 <- update_data(result_dir, filter_prolific_id, "steepness_noticed");
df2 <- update_data(result_dir, filter_quality, "quality");

```

