---
title: "results experiment 1"
output:
  html_document:
    df_print: paged
---

```{r, warning=FALSE, message=FALSE}
library(ggforce)
library(here)
library(tidyverse)
library(scales)
source(here("R", "utils.R"))
source(here("R", "utils-exp2.R"))
source(here("R", "utils-exp1.R"))

# debug_run <- TRUE
debug_run <- FALSE

exp.name = "experiment-wor(l)ds-of-toy-blocks"
fn <- paste(exp.name, "_tidy.rds", sep="");
RESULT_DIR <- ifelse(debug_run, here("data", "test-runs", "results", exp.name), 
                     here("data", "prolific", "results", exp.name))
PLOT_DIR <- paste(RESULT_DIR, "plots", sep=.Platform$file.sep);
if(!dir.exists(PLOT_DIR)){dir.create(PLOT_DIR, recursive=TRUE)}
data <- readRDS(paste(RESULT_DIR, fn, sep=.Platform$file.sep)) 

data.prior = data$test %>% filter(str_detect(trial_name, "multiple_slider")) %>% 
  select(-utterance, -custom_response)

data.production = data$test %>% filter(str_detect(trial_name, "fridge_view")) %>%
  select(-question, -utterance, -r_orig, -r_norm, -RT, -QUD, -n) # todo: utterance should not be there!
```

# Meta info about participants

```{r}
df <- data$info %>% group_by(prolific_id) %>% distinct()
p <- df %>%
  ggplot(aes(x=factor(0), y=timeSpent)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, width=0.2) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + sd(df$timeSpent),
                 color="yellow")) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + 2 * sd(df$timeSpent),
                 color="red")) +
  theme(legend.position="none") +
  labs(x = "", y="timeSpent in min.")
p

df %>% ungroup() %>%  group_by(prolific_id) %>%  summary()
```
# Reaction Times per stimulus

```{r}
df <- data$test %>% select(prolific_id, id, RT) %>% group_by(id) %>% distinct()

dat.RT = df %>% group_by(id) %>% summarise(mu=mean(RT), sd=sd(RT))
df <- left_join(df, dat.RT, by="id")
for(pa in seq(1,2)) {
  p <- df %>%  
    ggplot(aes(y=RT)) + geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(x=0, color=prolific_id), width = 0.1, alpha = 0.5) +
    geom_hline(aes(yintercept = mu), color="green", linetype="dashed") +
    theme(legend.position="right", axis.ticks.x = element_blank(), axis.text.x = element_blank()) +
    labs(x="", y="RT in ms", title="Reaction Times per stimulus") +
    facet_wrap_paginate(~id, nrow=2, ncol=5, page=pa) +
    scale_y_continuous(trans = log2_trans(),
                       breaks = trans_breaks("log2", function(x) 2^x),
                       labels = trans_format("log2", math_format(2^.x))) +
    theme(legend.position="none")
  print(p)
}
```

# General comments

```{r}
dat.comments = data$comments %>%
  filter(!is.na(comments) & comments!="")
print(dat.comments$comments)
```


# Color vision trials

```{r}
df <- data$color %>% group_by(prolific_id, id) %>%
  mutate(correct = expected == response)

df %>%
  ggplot(aes(x=id)) + geom_bar(position="dodge", aes(fill=correct)) +
  theme(axis.text.x = element_text(angle=90, hjust=1), legend.position = "top")

```

# Training trials

```{r, training-trials}
question_factors = c("ry", "r", "y", "none")
labels = c(ry="Red and Yellow", r="Red and ¬Yellow",
           y="¬Red and Yellow", none="¬Red and ¬Yellow")

df.train = data$train %>%
  mutate(response = as.numeric(response), response = response/100)%>%
  add_normed_exp1() %>%
  rename(response=r_norm) %>% select(-r_orig, -n) %>%
  filter(!is.na(response))

plotSliderRatings(df.train, question_factors, labels, cluster_by="r")
```



## Trials Prior-Elicitation 
Setup prior-eliciation tasks analysis

```{r, prior-elicitation-experiment}
data.prior.orig <- data.prior %>% mutate(response=r_orig) %>% select(-r_norm, -r_orig)
data.prior.norm <- data.prior %>% mutate(response=r_norm) %>% select(-r_norm, -r_orig)

data.prior.invalid = data.prior.norm %>% filter(is.na(response))
data.prior.norm <- data.prior.norm %>% filter(!is.na(response))

STIMULI <- data.prior.orig %>% pull(id) %>% unique()
N_participants <- data.prior.orig %>% pull(prolific_id) %>% unique() %>% length()

plot_dir <- paste(RESULT_DIR, "plots", sep=.Platform$file.sep);
dir.create(plot_dir, showWarnings = FALSE)
```

Use normalized data and check how many all-or-none answers were given since we
dont really want them, stimuli should have a little more uncertainty than that.

```{r, all-none-responses}
# Look for participants who gave all-or-none responses, if any,
# which stimuli were rated as all-or-none by how many participants?
df <- data.prior.norm %>% group_by(prolific_id, id) %>%
  summarise(n=max(response), .groups="drop_last") %>%
  filter(n==1) %>%
  mutate(id=factor(id, levels=STIMULI))

N0 <- N_participants - df %>% pull(prolific_id) %>% unique() %>% length()
df0 <- tibble(prolific_id=rep('--', N0), ratio_all_none=rep(0, N0))

df.participants_all_none <- df %>% group_by(prolific_id) %>%
  summarise(ratio_all_none = n() / length(STIMULI), .groups="drop_last") %>%
  mutate(prolific_id=as.character(prolific_id)) %>%
  bind_rows(df0)

if (nrow(df.participants_all_none) != 0){
  p <- df.participants_all_none %>% filter(prolific_id != "--") %>% 
        ggplot(aes(ratio_all_none)) +
        geom_bar(aes(x=prolific_id, y=ratio_all_none), stat="identity") +
        theme(axis.text.x = element_text(angle=90, hjust=1)) +
        labs(y="#participants", x="#all-none answers")
  print(p)

  df.stimuli_all_none <- df %>%
    ungroup() %>% group_by(id) %>%
    summarise(n_participants = n(), .groups="keep")

  p <- df.stimuli_all_none %>%
        ggplot(aes(x=id, y=n_participants)) + geom_bar(stat="identity") +
        scale_x_discrete(drop=FALSE) +
        theme(axis.text.x = element_text(angle=90, hjust=1)) +
        labs(y="#participants who gave all-none answer")
  print(p)
} else {
  print('nobody gave all-none responses (where sum=1)')
}
```


## Experimental data: probability tables
1. original responses
```{r, slider-ratings-prior}
questions = c("bg", "b", "g", "none")
labels <- c(bg="Blue and Green", b="Blue and ¬Green",
            g="¬Blue and Green", none="¬Blue and ¬Green")
plotSliderRatings(data.prior.orig, questions, labels, cluster_by="b")
``` 

2. normalized responses
```{r, slider-ratings-normed}
plotSliderRatings(data.prior.norm, questions, labels, cluster_by="b")
```


# Plot density of generated table means (normalized tables)

```{r, echo = FALSE, message=FALSE}
fn <- paste(RESULT_DIR, paste(exp.name, "_tables_mean.csv", sep=""),
            sep=.Platform$file.sep)
prob_means <- read_csv(fn) %>% add_probs() %>%
  pivot_longer(cols=c("b", "g", "bg", "none", starts_with("p_")), names_to = "key", values_to = "prob")
```


```{r, include=FALSE, eval=FALSE}
PlotDensity <- function(tables, p1, p2, save_as){
  df <- tables %>% filter(key==p1 | key == p2)
  p <- df %>%
    ggplot(aes(prob, fill=key)) +
    facet_wrap(~id) +
    geom_density(alpha=0.5)

  # ggsave(paste(TARGET_DIR, save_as, sep=.Platform$file.sep), p, width=6, height=4)
  return(p)
}
# Distributions of peoples' beliefs about probabilities across all stimuli

# tables %>% PlotDensity("p_a_given_c", "p_c_given_a", "p-conditionals.png")
# tables %>% PlotDensity("p_a", "p_c", "p-marginals.png")
# tables %>% PlotDensity("p_na_given_nc", "p_nc_given_na", "p-neg-conditionals.png")
# tables %>% PlotDensity("p_na", "p_nc", "p-neg-marginals.png")

PlotMeans <- function(df, id, by_id=FALSE) {
  p <- df %>% filter(key=="b" | key == "g" | key=="bg" | key=="none") %>%
    mutate(key=factor(key, levels=c("bg", "b", "g", "none"),
                      labels = c("blue and green", "only green", "only blue", "neither"))) %>% 
    ggplot(aes(x=factor(0), y=prob))
  
  if(by_id){
    p <- p + geom_bar(aes(fill=id), stat="identity", position="dodge") + labs(x="")
  } else {
    p <- p + geom_bar(stat="identity") + labs(x="", title=id)
  }
  
  p <- p +
    facet_wrap(~key, nrow=2, ncol=2) +
    theme(text = element_text(size=20))
  
  print(p)
  return(p)
}
```


```{r, table-means, echo=FALSE, eval=FALSE, include=FALSAE}
df.all <- prob_means %>% filter(!is.na(prob))
for(i in seq(1, length(STIMULI))) {
    df <- df.all %>% filter(id == STIMULI[[i]])
    if(df %>% nrow() != 0){
      PlotMeans(df, STIMULI[[i]]) 
    }
}
```
```{r}
expected.conjunction = c("if1_hh", "if1_lh", "if2_hl", "if2_ll",
                          "independent_hh", "independent_ll", "independent_hl",
                          "ind2");
expected.literal = c("independent_uh", "independent_ul");
expected.conditional = c("if1_uh", "if1_u-Lh", "if2_ul", "if2_u-Ll");
```

```{r, means-all}
df.all <- prob_means %>% filter(!is.na(prob))
p <- PlotMeans(df.all %>% filter(id %in% expected.conditional), "", by_id=TRUE) 
p <- PlotMeans(df.all %>% filter(id %in% expected.conjunction), "", by_id=TRUE) 
p <- PlotMeans(df.all %>% filter(id %in% expected.literal), "", by_id=TRUE) 

p <- PlotMeans(df.all %>%
               filter(id %in% c("if1_lh", "if1_u-Lh", "if1_uh", "if1_hh")), "",
               by_id=TRUE)

p <- PlotMeans(df.all %>% filter(id %in% c("if2_ll", "if2_u-Ll", "if2_ul",
                                           "if2_hl")), "", by_id=TRUE)

```



Prepare prior elicitation task data to compare with corresponding utterances

```{r}
df.prior_responses = data.prior.norm %>%
  select(-trial_name, -trial_number, -RT, -QUD, -group) %>% 
  pivot_wider(names_from = "question", values_from = "response") %>%
  add_probs()

prior_responses = df.prior_responses %>%
  pivot_longer(cols=c("b", "g", "bg", "none", starts_with("p_")),
               names_to="prob", values_to="val") %>%
  mutate(utterance=
           case_when(prob=="b" ~ "the blue block falls but the green block does not fall",
                     prob=="g" ~ "the green block falls but the blue block does not fall",
                     prob=="bg" ~ "both blocks fall",
                     prob=="none" ~ "neither block falls",
                     prob=="p_a" ~ "the blue block falls",
                     prob=="p_c" ~ "the green block falls",
                     prob=="p_na" ~ "the blue block does not fall",
                     prob=="p_nc" ~ "the green block does not fall",
                     prob=="p_c_given_a" ~ "if the blue block falls the green block falls",
                     prob=="p_a_given_c" ~ "if the green block falls the blue block falls",
                     prob=="p_nc_given_na" ~ "if the blue block does not fall the green block does not fall",
                     prob=="p_na_given_nc" ~ "if the green block does not fall the blue block does not fall",
                     prob=="p_likely_a" ~ "the blue block might fall",
                     prob=="p_likely_na" ~ "the blue block might not fall",
                     prob=="p_likely_c" ~ "the green block might fall",
                     prob=="p_likely_nc" ~ "the green block might not fall",
                     TRUE ~ ""
         ))

```


# Production Task
first standardize utterances (e.g.'but' and 'and' are treated identically).
All different utterances that were created are printed
```{r}
# summarize uttterances: use 'and' instead of 'but' everywhere
df.production <- data.production %>%
 mutate(across(where(is.factor), as.character),
        response= case_when(str_detect(response, "but") ~ str_replace(response, "but", "and"),
                            TRUE ~ response),
        utterance=response)

df.production = standardize_sentences(df.production)
df.production.empirical <- df.production %>%
  filter(id != "ind2") %>%
  group_by(id,response) %>%
  summarize(n=n(), .groups='drop_last') %>%
  mutate(N=sum(n), ratio=n/N) %>% 
  select(id, response, ratio) %>%
  add_column(predictor="empirical")
```

# Trials where conjunctions are expected

```{r, fig.width=12}
df.production.empirical %>%
  filter(id %in% expected.conjunction) %>%
  plotProductionTrials(PLOT_DIR, dat.prior_empirical=prior_responses)
# dat.model=test.model %>% filter(str_detect(id, "independent"))
```

# Trials where conditionals are expected

```{r, fig.width=12}
df.production.empirical %>%
  filter(id %in% expected.conditional) %>%
  plotProductionTrials(PLOT_DIR, dat.prior_empirical=prior_responses)
```

# Trials where literals/might are expected

```{r, fig.width=12}
df.production.empirical %>%
  filter(id %in% expected.literal) %>%
  plotProductionTrials(PLOT_DIR, dat.prior_empirical=prior_responses)
```




