---
title: "results joint experiment Pilot"
output:
  html_document:
    df_print: paged
---

Load data

```{r, warning=FALSE, message=FALSE}
library(here)
library(ggforce)
source(here("R", "joint_experiment", "analysis-utils.R"))
```

```{r}
df.prior = data.prior %>% select(-utterance, -custom_response)
df.production = data.production %>%
  select(-question, -r_orig, -r_norm, -QUD, -n) 
```

## Meta info about participants

```{r, meta, echo=FALSE}
df <- data$info %>% group_by(prolific_id) %>% distinct()
p <- df %>%
  ggplot(aes(x=factor(0), y=timeSpent)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16, width=0.2) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + sd(df$timeSpent),
                 color="yellow")) +
  geom_hline(aes(yintercept = mean(df$timeSpent) + 2 * sd(df$timeSpent),
                 color="red")) +
  theme(legend.position="none") +
  labs(x = "", y="timeSpent in min.")
p

df %>% ungroup() %>%  group_by(prolific_id) %>%  summary()
```

## Reaction Times per stimulus

```{r, reaction-times, echo=FALSE}
df <- df.production %>% select(prolific_id, id, RT) %>% group_by(id) %>%
  distinct()

dat.RT = df %>% group_by(id) %>% summarise(mu=mean(RT), sd=sd(RT), .groups="drop_last")
df <- left_join(df, dat.RT, by="id")
for(pa in seq(1,2)) {
  p <- df %>%  
    ggplot(aes(y=RT)) + geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(x=0, color=prolific_id), width = 0.1, alpha = 0.5) +
    geom_hline(aes(yintercept = mu), color="green", linetype="dashed") +
    theme(legend.position="right", axis.ticks.x = element_blank(), axis.text.x = element_blank()) +
    labs(x="", y="RT in ms", title="Reaction Times per stimulus") +
    facet_wrap_paginate(~id, nrow=2, ncol=5, page=pa) +
    scale_y_continuous(trans = log2_trans(),
                       breaks = trans_breaks("log2", function(x) 2^x),
                       labels = trans_format("log2", math_format(2^.x))) +
    theme(legend.position="none")
  print(p)
}
```

# General comments

```{r, echo=FALSE}
df.comments = data.comments %>% filter(!is.na(comments) & comments!="")
print(df.comments$comments)
```


# Color vision trials

```{r, color-vision, echo=FALSE}
df <- data.color %>% group_by(prolific_id, id) %>%
  mutate(correct = expected == response)

df %>%
  ggplot(aes(x=id)) + geom_bar(position="dodge", aes(fill=correct)) +
  theme(axis.text.x = element_text(angle=90, hjust=1), legend.position = "top")
pids.false_colors = df %>% filter(correct == FALSE)  %>% pull(prolific_id) 
```


```{r, slider-ratings-prior-normed, include=FALSE}
## Experimental data: probability tables with certain conditions
df.wide = TABLES.dep %>% filter(prolific_id %in% pids.false_colors) %>%
  group_by(prolific_id, id) %>%
  rename(bg=AC, b=`A-C`, g=`-AC`, none=`-A-C`) %>%
  select(bg, b, g, none, prolific_id, id) %>%
  rowid_to_column()
df.long = df.wide %>% 
  pivot_longer(cols=c(bg, b, g, none), names_to="question", values_to="response") 
df.long %>% 
  plotSliderRatings(questions.test, labels.test, cluster_by="bg")
```

# Plot Mean of cell values of generated tables (normalized tables)
Look at mean probability ratings dependent on the utterance expected in
respective production trial. 

```{r, plot-functions, echo=FALSE}
PlotMeans <- function(df, id, by_id=FALSE) {
  df.sorted = df %>% arrange(desc(means)) %>%
    mutate(label=case_when(question=="b" ~ labels.test[["b"]], 
                           question=="g" ~ labels.test[["g"]],
                           question=="bg" ~ labels.test[["bg"]],
                           question=="none" ~ labels.test[["none"]]))
  id.levels = df.sorted %>% pull(id) %>% unique()
  df.sorted = df.sorted %>%
    mutate(id=factor(id, levels=id.levels),
                     question=factor(question, levels=questions.test, labels=labels.test))
  p <- df.sorted %>%
    ggplot(aes(x=factor(0), y=means))
  
  if(by_id){
    p <- p + geom_bar(aes(fill=id), stat="identity", position="dodge") + labs(x="")
  } else {
    p <- p + geom_bar(stat="identity") + labs(x="", title=id)
  }
  
  p <- p +
    facet_wrap(~question, nrow=2, ncol=2) +
    theme(text = element_text(size=20))
  
  print(p)
  return(p)
}

expected.conjunction = c("if1_hh", "if1_lh", "if2_hl", "if2_ll",
                          "independent_hh", "independent_ll", "independent_hl",
                          "ind2");
expected.literal = c("independent_uh", "independent_ul");
expected.conditional = c("if1_uh", "if1_u-Lh", "if2_ul", "if2_u-Ll");

prob_means = data.prior.norm %>%
  select(-QUD, -RT, -group, -custom_response, -trial_name, -trial_number, -n, -utterance) %>%
  group_by(id, question) %>% 
  summarize(means=mean(response), sd=sd(response))

df.all <- prob_means %>% mutate(id=factor(id, levels = IDS))
```

```{r, plotting}
p <- PlotMeans(df.all %>%
                 filter(id %in% c("independent_ll", "independent_ul",
                                  "independent_hl", "independent_uh",
                                  "independent_hh")),
               "", by_id=TRUE)

p <- PlotMeans(df.all %>% filter(id %in% c("if1_lh", "if1_u-Lh", "if1_uh",
                                           "if1_hh")),
              "", by_id=TRUE)

p <- PlotMeans(df.all %>% filter(id %in% c("if2_ll", "if2_u-Ll", "if2_ul",
                                           "if2_hl")), "", by_id=TRUE)

```


## Production Task
Custom responses

```{r, custom-responses, fig.height = 10, fig.width = 10}
dat.custom = df.production %>% filter(!is.na(custom_response)) %>%
  group_by(prolific_id, id) %>% mutate(custom_response = tolower(custom_response))
total = df.production %>% select(prolific_id, id) %>% distinct()
print(paste('nb total custom responses given:', nrow(dat.custom), 'out of', nrow(total), 
            '(', nrow(dat.custom)/nrow(total), ')'))
df = dat.custom %>% rename(res.given=response, response=custom_response) %>%
  standardize_sentences() %>%
  mutate(newUtt = !response %in% standardized.sentences) %>%
  filter(newUtt) %>% rename(custom_response=response)

# dat.custom %>%
df %>% group_by(custom_response) %>% mutate(n=n()) %>% 
  ggplot(aes(x=n, y=fct_rev(fct_infreq(custom_response)))) +
  geom_bar(aes(fill=id), stat="identity", position=position_dodge()) +
  geom_text(aes(label=res.given), position = position_nudge(x=-0.5, y=-0.25)) +
  theme(text = element_text(size=20),
        axis.text.x=element_text(angle=45, vjust = 0.5),
        legend.position="none") +
  labs(y="#participants", x="custom response", title="(not creatable) custom responses") +
  theme_bw()
```


## Slider Ratings

```{r, prepare-exp1-data, echo=FALSE}
df.prior_responses = data.prior.norm %>%
  select(-custom_response, -QUD, -RT, -trial_number, -trial_name) %>% 
  pivot_wider(names_from = "question", values_from = "response") %>%
  add_probs()

prior_responses = df.prior_responses %>%
  pivot_longer(cols=c("b", "g", "bg", "none", starts_with("p_")),
               names_to="prob", values_to="val") %>%
  mutate(utterance=
           case_when(prob=="b" ~ standardized.sentences$b,
                     prob=="g" ~ standardized.sentences$g,
                     prob=="bg" ~ standardized.sentences$bg,
                     prob=="none" ~ standardized.sentences$none,
                     prob=="p_a" ~ "blue falls",
                     prob=="p_c" ~ "green falls",
                     prob=="p_na" ~ "blue does not fall",
                     prob=="p_nc" ~ "green does not fall",
                     prob=="p_c_given_a" ~ standardized.sentences$if_bg,
                     prob=="p_c_given_na" ~ standardized.sentences$if_nbg,
                     prob=="p_a_given_c" ~ standardized.sentences$if_gb,
                     prob=="p_a_given_nc" ~ standardized.sentences$if_ngb,
                     prob=="p_nc_given_a" ~ standardized.sentences$if_bng,
                     prob=="p_nc_given_na" ~ standardized.sentences$if_nbng,
                     prob=="p_na_given_c" ~ standardized.sentences$if_gnb,
                     prob=="p_na_given_nc" ~ standardized.sentences$if_ngnb,
                     prob=="p_likely_a" ~ "blue might fall",
                     prob=="p_likely_na" ~ "blue might not fall",
                     prob=="p_likely_c" ~ "green might fall",
                     prob=="p_likely_nc" ~ "green might not fall",
                     TRUE ~ NA_character_)
         )

exp1.human = prior_responses %>% select(-group, -n, -prob) %>%
  rename(human_exp1=val, stimulus_id=id)
exp2.human = df.production %>% select(prolific_id, id, response) %>%
  rename(stimulus_id=id, utterance=response) %>% add_column(human_exp2=1)
joint.human = left_join(exp1.human, exp2.human,
                        by=c("prolific_id", "stimulus_id", "utterance")) %>%
  filter(stimulus_id != "ind2")
```

```{r, message=FALSE, include = FALSE}
fn = paste(PLOT.dir, "by-participants", sep=SEP)
dir.create(fn)
plotSliderRatingsAndUtts(joint.human, fn)
```

```{r, slider-ratings-with-model}
# join-model-human1-data
exp2.model = data.model %>% select(-id) %>% 
  group_by(table_id,cn) %>%
  pivot_wider(names_from="response", values_from="probs", names_prefix="res.")
df.mapping = mapping_table_participant %>% select(-stimulus_src) %>% 
  group_by(table_id) %>%
  # separate(trial_id, into=c("stimulus_id", "prolific_id"), sep="--") %>%
  mutate(trial_id=list(trial_id)) %>% distinct()

#here id = stimulus_src
df = left_join(exp2.model, df.mapping, by=c("table_id")) %>%
  unnest(c(trial_id)) %>%
  separate(trial_id, into=c("stimulus_id", "prolific_id"), sep="--") %>% 
  pivot_longer(cols=starts_with("res."), names_to="response", values_to="probs",
               names_prefix="res.") %>% 
  rename(model_exp2=probs)

df.all_empirical = left_join(exp1.human, df, by=c("prolific_id", "stimulus_id", "response"))
df.all_joint = df.all_empirical %>% filter(!is.na(model_exp2))
saveRDS(df.all_joint, paste(RESULT.dir, "res-model-exp2-human-exp1.rds"))
# join-model-human2-data
exp2.human = df.production %>% select(prolific_id, id, response) %>%
  rename(stimulus_id=id, utterance=response)
df.plot = left_join(df.all_joint,
                    exp2.human %>% add_column(human_exp2=1) %>% rename(response=utterance))
```

Plot empirical ratings for each utterance against model predictions for respective utterances
for each participant separately.

```{r}
plotRatingsAndModel(df.plot, PLOT.dir)
```

Subjective probability ratings from Experiment 1 plotted dependent on utterance
participants chose.
  
```{r}
# 389 test trials (ind2 is not in here)
df = left_join(exp1.human,
               exp2.human %>% rename(response=utterance) %>%  
               add_column(human_exp2=1), by=c("stimulus_id", "prolific_id", "response")) %>%
  filter(!is.na(human_exp2)) %>% group_by(stimulus_id)

p <- df %>%
    ggplot(aes(y=response, x=human_exp1)) +
    geom_point(aes(color=stimulus_id)) +
    geom_vline(aes(xintercept=0.7), color="black", linetype='dashed') +
    # facet_wrap(~stimulus_id) +
    theme_bw(base_size = 20) +
    theme(text = element_text(size=20), legend.position="bottom") +
    labs(x="rated probability (prior elicitation)", y="response")

ggsave(paste(PLOT.dir, "prior-vs-utt.png", sep=SEP), p, width=15)#, height=18)
p
df %>% filter(human_exp1 >= 0.7) %>% nrow() / nrow(df)   
```

Literal meaning probabilities for responses that each participant gave in
respective trial, computed from participants' ratings in prior-elicitation task
(exp1) plotted separately for each participant. Stimuli are color coded.
When participants said 'X might/might not fall', it is checked whether the
corresponding rated probability is larger than 0.1, if yes, it is set to 1.

```{r}
# 389 test trials (ind2 from training is not in here)
df = left_join(exp1.human,
               exp2.human %>% rename(response=utterance) %>%  
               add_column(human_exp2=1), by=c("stimulus_id", "prolific_id", "response")) %>%
  filter(!is.na(human_exp2)) %>% group_by(stimulus_id) %>%
  mutate(human_exp1=case_when(str_detect(response, "might") & human_exp1>0.1 ~ 1,
                              TRUE ~ human_exp1))
p <- df %>%
    ggplot(aes(y=response, x=human_exp1)) +
    geom_point(aes(color=stimulus_id)) +
    geom_vline(aes(xintercept=0.7), color="black", linetype='dashed') +
    facet_wrap(~prolific_id) +
    theme_bw(base_size = 14) +
    theme(legend.position="none") +
    labs(x="rated probability (prior elicitation)", y="response")
ggsave(paste(PLOT.dir, "prior-vs-utt-per-proband.png", sep=SEP), p, width=15, height=18)
p
df %>% filter(human_exp1 >= 0.7) %>% nrow() / nrow(df) 

# Same plot for those participants who had at least one color-vision question wrong.
df.col = df %>% filter(prolific_id %in% pids.false_colors)
p <- df.col %>%
    ggplot(aes(y=response, x=human_exp1)) +
    geom_point(aes(color=stimulus_id)) +
    geom_vline(aes(xintercept=0.7), color="black", linetype='dashed') +
    facet_wrap(~prolific_id) +
    theme_bw(base_size = 14) +
    theme(legend.position="bottom") +
    labs(x="rated probability (prior elicitation)", y="response")
p
```



Quality of the data measured by the squared mean diff of a participant's prior rating to the 
mean of all participants' ratings for the respective probability table cell (of the respective stimulus), summed across all 4 cells for each participant. For each stimulus:
For each cell entry (P(A,C), P(A,¬C),..) $c$, compute mean rated probability:
$\mu_c = 1/N \cdot \sum_{i=1}^N x_{c,i}$.

That is, we get one value for each participant and stimulus. For participant i and 
a particular stimulus:
$\sum_{c=1}^{c=4}(x_{c,i} - \mu_c)^2$

jittered points are from those who gave at least one false response in color questions.
```{r}
df =  data.prior %>% responsesSquaredDiff2Mean() %>%
  mutate(stimulus_id=factor(stimulus_id))
df.colored = df %>%
  mutate(colored=prolific_id %in% pids.false_colors) %>% filter(colored==TRUE)

p = df %>%
  ggplot(aes(x=stimulus_id,  y=sum_sq_diff)) +
    geom_boxplot(aes(colour=stimulus_id), outlier.shape=NA) +
    geom_jitter(data=df.colored, aes(colour=stimulus_id), height=0, width=0.05, alpha=0.4, size=1.5) +
    theme_bw(base_size = 20) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5), legend.position="none")
p
ggsave(paste(PLOT.dir, "quality-check-sq-diff.png", sep=SEP), p, width=15, height=10)
```

For each stimulus, which response was created how often?

```{r, echo=FALSE, fig.height = 12}
df.production.means = df.production %>%
  group_by(id,response) %>%
  select(id, response, prolific_id) %>% summarize(n=n(), .groups="drop_last") %>%
  mutate(N=sum(n), ratio=n/N) %>% 
  select(id, response, ratio) %>%
  add_column(predictor="empirical")

df = df.production.means %>% 
  filter(ratio>0) %>%
  mutate(response=as_factor(response))

df %>% ggplot(aes(y=response, x=ratio)) +
  geom_bar(aes(fill=id), stat="identity") +
  theme_bw(base_size=18) +
  theme(legend.position = "none") +
  facet_wrap(~id)
```

Plot average model response per stimulus plotted against ratio of participants who
chose the respective response.
1.Trials where conjunctions are expected

```{r, fig.width=12, fig.height=14}
df.conj = bind_rows(df %>% filter(id %in% expected.conjunction),
                    model.avg %>% filter(id %in% expected.conjunction))
df.conj %>%
  plotProductionTrials(PLOT.dir)
```

2.Trials where conditionals are expected

```{r, fig.width=12, fig.height=15}
df.conditionals = bind_rows(df %>% filter(id %in% expected.conditional),
               model.avg %>% filter(id %in% expected.conditional))
df.conditionals %>%  
  plotProductionTrials(PLOT.dir)
```

3.Trials where literals/might are expected

```{r, fig.width=12, fig.height=15}
df.literals = bind_rows(df %>% filter(id %in% expected.literal),
               model.avg %>% filter(id %in% expected.literal))
df.literals %>%  
  plotProductionTrials(PLOT.dir)
```


Check whether many participants used all-or-none responses (0-0.5-1)

```{r, all-none-responses, include=FALSE}
# which stimuli were rated as all-or-none by how many participants?
df.all = data.prior.norm %>% group_by(prolific_id, id)

df <- df.all %>% summarise(n=max(response), .groups="drop_last")
df1 = df %>% mutate(id=factor(id, levels=c(IDS, "ind2"))) %>%
  filter(n==1)
ratio.one = nrow(df1) / nrow(df)
print(paste('ratio 1-0-0-0 answers', round(ratio.one,3)))

df05 = df.all %>% mutate(undecided=response>0.2 & response<=0.3) %>%
  filter(undecided==TRUE)
ratio.point5 = nrow(df05) / nrow(df.all)
print(paste('ratio roughly .25-.25-.25-.25 answers', round(ratio.point5, 3)))
```



